---
title: "00_text cleaning"
author: "Brenna Kelly"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(purrr)

```


- put all articles in a dataframe format
- grab quotes (and attribution) from the text
- 


```{r}

txt_files <- paste0("data/raw article text/", 
                    list.files("data/raw article text"))

read.delim(txt_files[6])

txt <- list()

for(i in 1:length(txt_files)) {
  
  txt[i] <- read.delim(txt_files[i])
  
}

txt_df <- as.data.frame(t(
  rbind(list.files("data/raw article text"),
                as.data.frame(t(txt)))
  ))

names(txt_df) <- c("file_name", "text")

test <- txt_df[13, "text"]

head(test)

write.csv(txt_df, "txt_df.csv", row.names = FALSE)

for(i in 1:length(txt_df)) {
  
  txt_df[i, "text"] <- unlist(txt_df[i, "text"])
  
}

unlist(txt_df[157, "text"])
txt_df[157, "file_name"]






sapply(txt_df[157, "text"], rbind)

lapply(unlist(txt_df[157, "text"]), rbind)

typeof(txt_df)

```


To get subset:  
- add year column
- get a 10% subset from each decade

```{r}

library(dplyr)
library(stringr)

txt_df$year <- str_sub(txt_df$file_name, start = 0, end = 4)
txt_df$decade <- paste0(str_sub(txt_df$year, start = 0, end = 3), "0s")

table(txt_df$year)
table(txt_df$decade)

# the proportion from each decade
# - note: combine 2010s and 2020s
p_80s <- ceiling(0.1 * table(txt_df$decade)["1980s"])
p_90s <- ceiling(0.1 * table(txt_df$decade)["1990s"])
p_20s <- ceiling(0.1 * table(txt_df$decade)["2000s"])
p_1020s <- ceiling(0.1 * table(txt_df$decade)["2010s"]) + 
  ceiling(0.1 * table(txt_df$decade)["2020s"])

# decade for random subset
txt_df$decade_subset <- ifelse(txt_df$decade %in% c("2010s", "2020s"),
                               "2010s, 2020s", txt_df$decade)
table(txt_df$decade_subset)
round(prop.table(table(txt_df$decade_subset)), 2)

# select random subset, stratified by decade of sampling (not using what's above)
set.seed("3006")

txt_subset <- txt_df |>
  group_by(decade_subset) |>
  sample_frac(size = 0.1)

lapply(txt_subset$file_name, )



```

Putting it into a format with one "line" of text per row, for candidate tone identification:

```{r}


txt_reformat <- list()

for(i in 1:nrow(txt_subset)) {
  
  test <- map_df(txt_subset$text[i], ~as.data.frame(.))
  
  test <- map_df(test, ~as.data.frame(.))
  print(str(test))
  
  # adding other columns
  test$file_name <- txt_subset$file_name[i]
  test$year <- txt_subset$year[i]
  test$decade <- txt_subset$decade[i]
  
  txt_reformat[[i]] <- test
  
}

txt_reformat <- do.call(rbind, txt_reformat)

txt_reformat$file_name <- do.call(rbind, txt_reformat$file_name)

str(txt_reformat)

write.csv(txt_reformat, "txt_subsample.csv", row.names = FALSE)

```

list of practice candidate tones

```{r}

library(tidyr)

tones <- read.csv("../practice_candidate tones.csv")

tones <- tones |>
  filter(!TONES %in% c("", " ", "but", "standpoint", "na", "ve")) |>
  mutate(TONES = tolower(TONES))

unique_tones <- tones |>
  separate_rows(TONES) |>
  distinct()

unique_tones <- data.frame(tone = do.call(paste, c(as.data.frame(t(unique_tones)), sep = ", ")))

unique_tones

```



- can I apply the model to an observation in txt_df?
- can I apply the model to an item in list? (each line in an observation in txt_df)

```{r}



```





